---
title: Correlation
subtitle: Definition, types of correlation, scatter diagram, Karl Pearson's coefficient of correlation (linear correlation)
author:   
  - Deependra Dhakal 
institute:   
  - Assistant Professor
  - Agriculture and Forestry University
  - \url{https://rookie.rbind.io}
output:   
  beamer_presentation:  
    incremental: false  
    theme: "Frankfurt"  
    colortheme: "beaver"  
    toc: true   
    slide_level: 2
    keep_tex: true
    includes:
      in_header: beamer_header_template.tex
classoption: "aspectratio=169"
header-includes: 
- \captionsetup[table]{skip=0pt}
- \captionsetup[table]{belowskip=-2pt}
- \setbeamerfont{caption}{size=\tiny}
- \setbeamertemplate{footline}[page number]
- \AtBeginSection{}
bibliography: [./bibliographies.bib]
---

```{r setup, include=FALSE}
library(knitr)
require(tidyverse)
require(gridExtra)
set.seed(453)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, echo = FALSE, 
                  message = FALSE, warning = FALSE,
                  out.width = "45%")
options(knitr.table.format = "latex")
options(knitr.kable.NA = "", digits = 2)
options(kableExtra.latex.load_packages = FALSE)

# set penalty for using scientific notation at 10, and number of digits to use in notation to 10
options(scipen = 30, digits = 15)

theme_set(theme_bw())
```

# Correlation

## Meaning and definition

- Suppose we have a sample of n pairs for which each pair represents the measurement of two variables, $X$ and $Y$. If a scatterplot of $Y$ versus $X$ shows a general linear trend, then it is natural to try to describe the strength of the linear association.
- The systematic interrelationship between the two continuous related variables say, X and Y is termed as correlation. When only two variables are involved, the correlation is called simple correlation. If more than two variables are involved, the correlation is said to be multiple correlation.
- When the variables move in the same direction, i.e., increase in one variable causes and increase in other variable and _vice versa_, such type of correlation is called positive/direct correlation. In general, grain yield of wheat and the number of grains per spike are positively correlated. 
- By analogy, negative correlation is said to occur when increase in one variable is followed by decrease in other. For example, grain yield of wheat and severity of disease in the field are negatively correlated.

##

\begin{columns}[T, onlytextwidth]
\column{0.5\textwidth}
\small Anscombe's Quartet

```{r anscombe_quartet}
anscombe <- datasets::anscombe

anscombe %>%
  mutate(SN = row_number()) %>% 
  relocate(SN, .before = 1) %>% 
  knitr::kable(booktabs = TRUE, caption = "Anscombes quartet is a set of 4 ($x,y$) data sets that were published by Francis Anscombe in a 1973 paper Graphs in statistical analysis.", linesep = "") %>% 
  kableExtra::kable_styling(latex_options = "striped", font_size = 5) %>% 
  kableExtra::column_spec(column = 1:8, width = c(rep("1.5em", 4), rep("2em", 4)))

```

\column{0.5\textwidth}
\small Rothamsted Oats

```{r rothamsted_oats}
rothamsted_oat <- agridat::rothamsted.oats

mutate(head(rothamsted_oat[, c("grain", "straw")], n = 6),
       across(everything(), ~as.character(.))) %>% 
  bind_rows(tibble(grain = rep("$\\dots$", 3), straw = rep("$\\dots$", 3))) %>% 
  bind_rows(mutate(tail(rothamsted_oat[, c("grain", "straw")], n = 6),
       across(everything(), ~as.character(.)))) %>%
  mutate(SN = row_number()) %>% 
  relocate(SN, .before = 1) %>% 
  knitr::kable(booktabs = TRUE, caption = "A dataset from RCB experiment, carried out at Rothamsted facility, of oats taking measurements on straw and grain with 9 fertilizer treatments.", escape = FALSE, linesep = "") %>% 
  kableExtra::kable_styling(font_size = 5, latex_options = "striped") %>% 
  kableExtra::column_spec(column = 1:2, width = c("5em", "5em"))

```

\end{columns}

\footnotesize
- We take the two datasets (above) and show (graphically) following types of correlation in scatterplot:
  \begin{itemize}
  \scriptsize
  \item Positive correlation
  \item Negative correlation
  \item Dinosaur correlation
  \end{itemize}

## Scatterplot diagram

```{r correlation-scatter-plot, fig.align='center', fig.cap="Association between variables in two datasets (Anscombe's Quartet: Left; Rothamsted experiment: Right)", out.width="98%", fig.height=3, fig.width=6}
require(cowplot)
p1 <- ggplot(anscombe) +
  geom_point(aes(x1, y1), color = "darkorange", size = 1.5) +
  scale_x_continuous(breaks = seq(0,20,2)) +
  scale_y_continuous(breaks = seq(0,12,2)) +
  expand_limits(x = 0, y = 0) +
  labs(x = "x1", y = "y1",
       title = "Dataset 1" ) +
  theme_bw()
# p1

rothamsted_oat_gg <- ggplot(rothamsted_oat) +
  geom_point(aes(straw, grain), color = "darkblue", size = 1.5) +
  # scale_x_continuous(breaks = seq(0,20,2)) +
  # scale_y_continuous(breaks = seq(0,12,2)) +
  # expand_limits(x = 0, y = 0) +
  labs(x = "x1", y = "y1",
       title = "Relationship between Straw and Yield of Oat" ) +
  theme_bw()
# rothamsted_oat_gg

plot_grid(p1, rothamsted_oat_gg, align = "h",
          ncol = 2, rel_widths = c(1/6, 1/4))
```

##

```{r}
# Mean
sapply(1:8, function(x) mean(anscombe[ , x]))
## [1] 9.000000 9.000000 9.000000 9.000000 7.500909 7.500909 7.500000 7.500909
# Variance
sapply(1:8, function(x) var(anscombe[ , x]))
## [1] 11.000000 11.000000 11.000000 11.000000  4.127269  4.127629  4.122620
## [8]  4.123249
# Coorelation
sapply(1:4, function(x) cor(anscombe[ , x], anscombe[ , x+4]))
```

##

```{r}
skimr::skim(anscombe) %>%
  knitr::kable(booktabs = TRUE)
```


## Mathematical representation of correlation


## Karl pearson's coefficient of correlation (linear correlation)

## Correlation coefficient for bivariate frequency distribution