---
title: Descriptive measures
author: Deependra Dhakal
date: August, 2019
institute: |
  | \textit{ddhakal.rookie@gmail.com}
  | \url{https://rookie.rbind.io}
output: 
  binb::metropolis:
    includes:
      in_header: beamer_header_template.tex
classoption: "aspectratio=169" # "compressed"
fontsize: 11pt
bibliography: [./../bibliographies.bib]
---

```{r,setup, include=FALSE}
library(knitr)
require(tidyverse)
set.seed(453)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, echo = FALSE, 
                  message = FALSE, warning = FALSE,
                  out.width = "45%", cache = TRUE)
options(knitr.table.format = "latex")
options(knitr.kable.NA = "", digits = 2)
options(kableExtra.latex.load_packages = FALSE)
theme_set(theme_bw())
```

# Quantitative data

In a mass, quantitative data show following characteristics:

1. They show a tendency to concentrate at certain values, usually somewhere in centre of the distribution. Measures of this tendency are called measures of central tendency or averages.

2. The data vary about a measure of central tendency and these measures of deviation are called measures of *variation* or *dispersion*.

3. The data in a frequency distribution may fall into symmetrical or asymmetrical patterns. The measures of the direction and degree of asymmetry are called measures of *skewness*.

4. Polygons of frequency distributions exhibit flatness or peakedness of frequency curves. The measures of peakedness or flatness of the frequency curves are called measures of *kurtosis*.

## Frequency distribution

When observations of either discrete or continuous nature are available on a single characteristic of a large number of individuals, often it becomes necessary to condense the data as far as possible without losing any information. Let us consider the marks in a subject obtained by 300 candidates selected at random among those appearing a certain examination.

```{r uniform-dist-std, echo=FALSE}
students_marks <- runif(300, 32, 99) %>% enframe(name = "student", value = "marks")

students_marks$marks %>% 
  matrix(nrow = 30) %>% 
  as_tibble() %>% 
  rename()
```

